### 2014-07-22 MS.tex
typesetting

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{hyperref}

% line numbers
\usepackage{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% rotating package for sideways tables
%\usepackage{rotating}

% If you wish to include algorithms, please use one of the packages below. Also, please see the algorithm section of our LaTeX guidelines (http://www.plosone.org/static/latexGuidelines) for important information about required formatting.
%\usepackage{algorithmic}
%\usepackage{algorithmicx}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

\author{}
\date{}
\pagestyle{myheadings}

\begin{document}
% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Diversity dynamics in Nymphalidae butterflies: Effect of
phylogenetic uncertainty on diversification rate shift estimates}
}
% Insert Author names, affiliations and corresponding author email.
\\
Carlos Pe\~na$^{1,\ast}$, 
Marianne Espeland$^{2}$
\\
\bf{1} Carlos Pe\~na Laboratory of Genetics, Department of Biology, University of Turku, Turku
\\
\bf{2} Museum of Comparative Zoology and Department of Organismic and
Evolutionary Biology, Harvard University, Cambridge, USA. Email:
\href{mailto:marianne.espeland@gmail.com}{marianne.espeland@gmail.com}.
Current address: McGuire Center for Lepidoptera and Biodiversity,
Florida Museum of Natural History, University of Florida, Gainesville,
FL, US
\\
$\ast$ E-mail: Corresponding mycalesis@gmail.com
\end{flushleft}

\subsection{Abstract}\label{abstract}

The species rich butterfly family Nymphalidae has been used to study
evolutionary interactions between plants and insects. Theories of
insect-hostplant dynamics predict accelerated diversification due to key
innovations. We investigated whether phylogenetic uncertainty affects a
commonly used method (MEDUSA, modelling evolutionary diversity using
stepwise AIC) for estimating shifts in diversification rates in
lineages, by extending the method across a random sample of trees from
the posterior distribution of a Bayesian run. We found that phylogenetic
uncertainty greatly affects diversification rate estimates. Different
trees produced diversification rates ranging from high values to almost
zero for the same clade, and both significant rate increase and decrease
in some clades. Only four out of 18 significant shifts found on the
maximum clade credibility tree were consistent across most of the
sampled trees. Among these, we found accelerated diversification for
Ithomiini butterflies. We used the binary speciation and extinction
model (BiSSE) and found that a hostplant shift to Solanaceae is
correlated with increased net diversification rates in Ithomiini,
congruent with the diffuse cospeciation hypothesis. Our results show
that taking phylogenetic uncertainty into account when estimating net
diversification rate shifts is of great importance, and relying on the
maximum clade credibility tree alone can potentially give erroneous
results.

\textbf{Keywords:} diversification analysis, MEDUSA, BiSSE, speciation
rate, insect-hostplant dynamics

\subsection{Introduction}\label{introduction}

Hostplant shifts have been invoked to be responsible for a great part of
the biodiversity of herbivorous insects {[}1{]}. The study of the
evolution of hostplant use has spawned several theories explaining the
evolutionary interactions between plants and insects {[}2{]}. The
``escape-and-radiate'' hypothesis {[}3{]}, the ``oscillation
hypothesis'' {[}4,5{]} or ``diffuse cospeciation'' {[}2{]}.

The butterfly family Nymphalidae has been an important taxon for
developing some of the mentioned hypotheses. Nymphalidae contains around
6000 species {[}6{]}, and is the largest family within the true
butterflies. The family most likely originated around 94 MYA in the mid
Cretaceous. Diversification of the group began in the Late Cretaceous
and most major radiations (current subfamilies) appeared shortly after
the Cretaceous-Paleogene (K-Pg) boundary {[}7{]}. Several studies have
used time-calibrated phylogenies and diversification models to
reconstruct the evolutionary history of the group to identify patterns
of accelerated or decelerated diversification of some Nymphalidae clades
{[}7--10{]}. For example, it has been suggested that climate change in
the Oligocene and the subsequent diversification of grasses has led to
diversification of the subfamily Satyrinae {[}11{]} due to the abundance
of grasses over extensive geographic areas (``resource
abundance-dependent diversity dynamics'' hypothesis). Reference 9 found
increased net diversification rates in some Nymphalidae lineages after a
major hostplant shift, which appears to be in agreement with Reference 3
``escape-and-radiate'' model of diversification.

Although it has been suggested that part of the great diversity of
Nymphalidae butterflies is a result of hostplant-insect dynamics, it is
necessary to use modern techniques to investigate whether the
diversification patterns of Nymphalidae are in agreement with the
theoretical predictions. It is necessary to test whether the overall
diversification pattern of Nymphalidae is congruent with events of
sudden diversification bursts due to hostplant shift, climatic events or
shifts to closely related hostplants {[}5,12{]}.

In this study, we used a time-calibrated genus-level phylogenetic
hypothesis for Nymphalidae butterflies {[}10{]} to investigate patterns
of diversification. We applied MEDUSA {[}13,14{]}, a recently developed
statistical method, to study the diversification pattern of Nymphalidae
butterflies. MEDUSA fits likelihood models of diversification into a
time-calibrated tree and tests whether allowing increases or decreases
in speciation and extinction rates within the tree produces better fit
of the models. MEDUSA is able to take into account unsampled extant
species diversity during model fitting and it is normally applied to the
maximum clade credibility phylogenetic tree. Particularly, we wanted to
study the effects of phylogenetic uncertainty and by using the extended
MEDUSA method called MultiMEDUSA {[}13{]}. We also tested whether
hostplant association dynamics can explain the diversification patterns
of component Nymphalidae lineages by testing whether character states of
hostplant use affected the diversification pattern of those lineages
employing the method BiSSE as implemented in the R package
\texttt{diversitree} {[}15{]}.

\subsection{Methods}\label{methods}

\subsubsection{Data}\label{data}

For analyses, we used the phylogenetic trees from the study of Reference
10 that were generated using DNA sequence data from 10 gene regions for
398 of the 540 valid genera in Nymphalidae. We employed Reference 10
maximum clade credibility tree (MCC tree) (Fig. 1) as well as a random
sample of 1000 trees from their BEAST run after burnin. Their original
BEAST run was for 40 million generations. We used a burnin of 25 million
generations and took a random sample of 1000 trees using Burntrees
v.0.1.9 \url{http://www.abc.se/~nylander/} (supp. mat. 5) in order to
correct for phylogenetic uncertainty when performing the diversification
analyses.

Species richness data for Nymphalidae genera were compiled from several
sources including the specialist-curated lists on
\url{http://tolweb.org}, Reference 16 and curated lists of Global
Butterfly Names project (\url{http://www.ucl.ac.uk/taxome/gbn/}). We
assigned the species numbers of genera not included in the phylogeny to
the closest related genus that was included in Reference 10 study
according to available phylogenetic studies {[}17--28{]}, taxonomical
classification and morphological resemblance when no phylogenies were
available (supp. mat. 3).

Hostplant data for Nymphalidae species were compiled from several
sources including Reference 29, HOSTS database
(\url{http://bit.ly/YI7nwW}), Reference 30, Reference 31, Reference 32
and others (supp. mat. 15--16) for a total of 6586 hostplant records,
including 428 Nymphalidae genera and 143 plant families and 1070 plant
genera. It was not possible to find any hostplant data for 35 butterfly
genera.

\subsubsection{Analyses of
Diversification}\label{analyses-of-diversification}

We used the statistical software R version 3.0.1 {[}33{]} in combination
with the APE {[}34{]}, GEIGER {[}35{]}, MEDUSA {[}13{]} and
\texttt{diversitree} {[}15{]} packages along with our own scripts to
perform the analyses (included as supplementary materials 2, 6, 10). All
analyses were run on the 1000 random trees from Reference 10, on the
maximum clade credibility tree derived from these and the MCC tree from
Reference 10.

\subsubsection{Detecting diversification shifts on phylogenetic
trees}\label{detecting-diversification-shifts-on-phylogenetic-trees}

Patterns of diversification in Nymphalidae were analyzed by using MEDUSA
version 093.4.33 {[}13{]} on the maximum clade credibility tree from
Reference 10. MEDUSA fits alternative birth-death likelihood models to a
phylogenetic tree in order to estimate changes in net diversification
rates along branches. MEDUSA estimates likelihood and AIC scores for the
simplest birth-death model, with two parameters, the rates \texttt{r}:
net diversification \texttt{=} speciation (\texttt{b}) \texttt{-}
extinction (\texttt{d}) and $\varepsilon$: relative extinction
\texttt{= d/b}. The AIC scores of the two-parameter model are then
compared with incrementally more complex models until the addition of
parameters do not improve the AIC scores beyond a cutoff value. MEDUSA
finds the likelihood of the models after taking into account branch
lengths and number of species per lineage {[}13{]}. Most studies using
MEDUSA only run the method on a single tree, usually the maximum clade
credibility tree, which makes the assumption that this tree is correct.
We wanted to study the effects of phylogenetic uncertainty on estimation
of net diversification rate shifts and therefore selected 1000 random
genus-level trees from the posterior distribution (MultiMEDUSA, supp.
mat. 5). We calculated a new MCC tree from the selection of trees, ran
MEDUSA and MultiMEDUSA on the selection of 1000 trees and summarized the
estimated changes in net diversification rates for nodes across all
trees. Patterns of change in net diversification rates are considered
significant if they are found at the same node in at least 90\% of the
trees. We also expected to find similar values of net diversification
(\texttt{r = b - d}) and relative extinction rate ($\varepsilon$
\texttt{= d/b}) values across the 1000 trees for the nodes where changes
in diversification tempo occurs. We used the AICc threshold of 7.8
units, as estimated by MEDUSA, as the limit for a significantly better
fit to select among increasingly complex alternative models.

\subsubsection{Estimation of trait-dependent speciation
rates}\label{estimation-of-trait-dependent-speciation-rates}

As the MEDUSA and MultiMEDUSA approaches estimated an increase in net
diversification in the clade Ithomiini, we tested whether this pattern
can be explained by increase in birth-rate due to hostplant use and
performed analyses with the ``binary state speciation and extinction''
{[}36{]} Bayesian approach as implemented in the R package
\texttt{diversitree} {[}15{]}. MuSSE {[}15{]} is designed to examine the
joint effects of two or more traits on speciation. Because most of
Nymphalidae butterflies are restricted to use one plant family as
hostplant, the character states can be coded as presence/absence, for
which the BiSSE analysis is better suited. BiSSE was designed to test
whether a binary character state has had any effect on increased net
diversification rate for a clade {[}36{]}. We used our compiled data of
hostplant use to produce a binary dataset for the character ``feeding on
the plant family Solanaceae'' with two states (absence = 0; presence =
1) (supp. mat. 17), which is the main hostplants of the diverse
Ithomiini butterflies and closest relatives {[}37{]}. Other hostplant
shifts were not tested for effect on net diversification rates due to
the low support for diversification shifts found in the MEDUSA analyses.
We analyzed the data using BiSSE employing the Markov Chain Monte Carlo
algorithm on the maximum clade credibility tree, taking into account
missing taxa by using the parameter ``sampling factor''
(\texttt{sampling.f}) in \texttt{diversitree}. We also used constrained
analyses forcing no effect of hostplant use on diversification and used
likelihood ratio tests to find out whether the hypothesis of effect on
diversification has a significantly better likelihood than the null
hypothesis (no effect). The analyses were run across a sample of 250
trees from the posterior distribution. The records of \emph{Vanessa} and
\emph{Hypanartia} feeding on Solanaceae {[}30,38{]} might be incorrect
as it is unlikely that these species can be feeding on this plant
family. We tested whether coding these two genera as not feeding on
Solanaceae affected our results.

\subsection{Results}\label{results}

\subsubsection{Detecting diversification shifts on the maximum clade
credibility
tree}\label{detecting-diversification-shifts-on-the-maximum-clade-credibility-tree}

The MEDUSA analysis on the MCC tree in combination with richness data
estimated 18 changes in the tempo of diversification in Nymphalidae
history (Fig. 1; Table 1). The corrected AICc acceptance threshold for
adding subsequent piecewise birth-death processes to the overall model
was set to 7.8 units, as prescribed by MEDUSA. In all MEDUSA analyses,
the maximum number of inferred diversification splits in all trees was
26. The background net diversification rate for Nymphalidae was
estimated as \texttt{r = 0.092} lineages per Million of years and the
AICc score for the best fit model was \texttt{5090.5} (Table 1). MEDUSA
also estimated that the basic constant birth--death model was not a
better explanation for our data (AICc \texttt{= 5449.3}).

Some of the 18 changes in diversification correspond to rate increases
in very species-rich genera: \emph{Ypthima} (\texttt{r = 0.311}),
\emph{Charaxes} (\texttt{r = 0.291}), \emph{Callicore} +
\emph{Diaethria} (\texttt{r = 0.220}), \emph{Pedaliodes}
(\texttt{r = 0.196}) and \emph{Taenaris} (\texttt{r = 0.234}). We found
rate increases for other clades as well such as: Mycalesina
(\texttt{r = 0.191}), Oleriina + Ithomiina + Napeogenina + Dircennina +
Godyridina (\texttt{r = 0.187}), Satyrini (\texttt{r = 0.116}),
Phyciodina in part (\texttt{r = 0.241}) and Satyrina
(\texttt{r = 0.221}), \emph{Coenonympha} (\texttt{r = 0.209}),
\emph{Caeruleuptychia} + \emph{Magneuptychia} (\texttt{r = 0.312}) and
\emph{Taenaris} (\texttt{r = 0.312}). We also found decreases in net
diversification rates for Limenitidinae + Heliconiinae
(\texttt{r = 0.0541}), part of Danaini (\texttt{r = 0.0423}),
Pseudergolinae (\texttt{r = 0.024}) and Coenonymphina
(\texttt{r = 0.065}) (Table 1).

\subsubsection{Phylogenetic uncertainty in the MultiMEDUSA
approach}\label{phylogenetic-uncertainty-in-the-multimedusa-approach}

We tested MEDUSA to find out whether taking into account the
phylogenetic signal from a random sample of 1000 trees from the
posterior distribution can return similar estimates of diversification
to the values obtained from the MCC tree. We run MultiMEDUSA on a random
sample of 1000 trees (supp. mat. 5) from the posterior distribution and
compared the results with a MEDUSA analysis on the MCC tree derived from
this sample (supp. mat. 9).

We found that the analysis by MultiMEDUSA on the 1000 trees estimated
lower median net diversification rates for the diversification shifts
found by MEDUSA on the MCC tree from Reference 10 (Table 2). Although
the diversification pattern found by MEDUSA and MultiMEDUSA was the
same, the latter consistently estimated lower rates. Thus, the shifts
recovered with low net diversification rate on the MCC were recovered
with negative net diversification rate by MultiMEDUSA. The background
diversification and all shifts found by MEDUSA on the 1000 trees are
provided as an R object in supp. mat. 7).

We also compared the results from MultiMEDUSA (derived from the sample
of 1000 trees) with the splits found by MEDUSA on the MCC tree derived
from this random sample. In the summary statistics, MultiMEDUSA reports
the frequency of the diversification shifts found in the trees
(parameter \texttt{sum.prop}). Thus, if a node is found in only half of
the 1000 trees, but the phylogenetic signal was strong enough to be
picked up by MEDUSA and a node shift was found most of the time, then
the \texttt{sum.prop} should be close to 1. For example the node 625
consists of the clade \texttt{Charaxes} + \texttt{Polyura} and it was
found in only 256 trees, however MultiMEDUSA was consistenly able to
find a diversification shift for that node and the \texttt{sum.prop}
value is 0.996.

We found different rates for diversification shifts even for the nodes
that appear in most of the sample of 1000 trees (Table 2) as many of the
nodes also found in the MCC tree had a probability of being recovered
higher than 0.90.

There were four diversification shifts found with more than 90\% of
probability in the trees from the random sample (Table 2): (i) net
diversification rate increase in the genus \emph{Ypthima}; (ii) net
diversification rate increase in the genus \emph{Charaxes}; (iii) rate
increase in Ithomiini subtribes Oleriina + Ithomiina + Napeogenina +
Dircennina + Godyridina; and (iv) rate increase in \emph{Callicore} +
\emph{Diaethria}.

MultiMEDUSA provided mean and standard deviation statistics for the
diversification values found on the shifts on the 1000 trees (supp. mat.
13--14), and found that some of the changes in net diversification rate
values had great variation across the posterior distribution of trees. A
boxplot of the net diversification rate values estimated for the clades
that appear in the MCC tree shows that some shifts are estimated as
increased or slowed diversification pace depending on the tree used for
analysis (Fig. 2). This variation is especially wide for the clade
formed by the genera \emph{Magneuptychia} and \emph{Caeruleuptychia}
because MEDUSA estimated diversification values from six times the
background net diversification rate (\texttt{r = 0.5234}) to almost zero
(\texttt{r = 1.22e-01}). The rates for \emph{Taenaris} were between 0.14
and 0.44 (mean value 0.25). Similar degrees of variation were found in
the nodes for \emph{Ypthima}, \emph{Charaxes} and \emph{Coenonympha}
(Fig. 2). The net diversification rates estimates for the clades
(Oleriina + Ithomiina + Napeogenina + Dircennina + Godyridina),
Limenitidinae + Heliconiinae and Pseudergolinae are relatively
consistent across the 1000 trees (Fig. 2).

It is also evident that not all the diversification shifts estimated on
the MCC tree are consistently recovered in most of the 1000 trees. Some
of the splits in the MCC tree are recovered in very few trees, for
example the split for the clade Satyrini (Euptychiina + Pronophilina +
Satyrina + Maniolina) is recovered with a probability of 18\% (Fig. 3).

\subsubsection{Estimation of trait-dependent speciation
rates}\label{estimation-of-trait-dependent-speciation-rates-1}

The MEDUSA analyses taking into account phylogenetic uncertainty
estimated a net diversification rate increase in part of the clade
Ithomiini across more than 95\% of the trees. Our BiSSE analysis found a
positive effect of the character state ``feeding on Solanaceae'' on the
net net diversification rate on part of Ithomiini (Oleriina + Ithomiina
+ Napeogenina + Dircennina + Godyridina) (Fig. 4). The Markov Chain
Monte Carlo algorithm was run for 10000 generations discarding the first
7500 as burnin. The estimated mean net diversification rate for taxa
that do not feed on Solanaceae was \texttt{r = 0.11} while the net
diversification rate for the Solanaceae feeders was \texttt{r = 0.16}
(see Fig. S1 for a boxplot of speciation and extinction values for the
95\% credibility intervals). The same analysis considering
\emph{Vanessa} and \emph{Hypanartia} as non-Solanaceae feeders due to
dubious records produced the same pattern and net diversification rates
(Fig. S2). Therefore, the rest of the analyses were performed assuming
these two genera as Solanaceae feeders. We constrained the BiSSE
likelihood model to force equal rates of speciation for both character
states in order to test whether the model of different speciation rates
is a significantly better explanation for the data. A likelihood ratio
test found that the model for increased net diversification rate for
nymphalids feeding on Solanaceae is a significantly better explanation
than this character state having no effect on diversification
(\texttt{p \textless{} 0.001}) (Table 3; character states available in
supp. mat. 17, code in supp. mat. 18, and mcmc run in supp. mat. 19). We
combined the post-burnin mcmc generations from running BiSSE on 250
trees from the posterior distribution and found the same pattern as the
BiSSE analysis on the maximum clade credibility tree (combined mcmc run
in supp. mat. 20; profiles plot of speciation rates in Fig. S3; boxplot
of 95\% credibility intervals in Fig. S4). A BiSSE analysis to test
whether the trait ``feeding on Apocynaceae'' had any effect on increased
net diversification rates found similar speciation rates for lineages
feeding on Apocynaceae and other plants (Fig. S5). It has been shown
that BiSSE performs poorly under certain conditions {[}39{]}. However,
our data has adequate number of taxa under analysis (more than 300
tips), adequate speciation bias (between 1.5x and 2.0x), character state
bias (around 8x) and extinction bias (around 4x) for the analysis of
Solanaceae hostplants. Thus, BiSSE is expected to produce robust results
{[}39{]}.

\subsection{Discussion}\label{discussion}

\subsubsection{Effects of phylogenetic uncertainty on the performance of
MEDUSA}\label{effects-of-phylogenetic-uncertainty-on-the-performance-of-medusa}

The MEDUSA method has been used to infer changes in net diversification
rates along a phylogenetic tree. Since its publication {[}13{]} the
results of using MEDUSA on a single tree, the maximum clade credibility
tree, have been used for generation of hypotheses and discussion
{[}7,40,41{]}. However, we found that MEDUSA estimated different
diversification shifts and different rates of diversification for
certain lineages when phylogenetic uncertainty was taken into account by
using MEDUSA on a random sample of trees from the posterior distribution
of a Bayesian run. We found that some diversification splits, estimated
on the Nymphalidae maximum clade credibility tree, were found with a
very low probability in a random sample of 1000 trees from the posterior
distribution (Fig. 3, Table 2). We also found that, even though MEDUSA
could estimate the same diversification splits on two or more trees, the
estimated net diversification rates could vary widely (Fig. 2). For
example, in our Nymphalidae trees, we found that the split for
\emph{Magneuptychia} and \emph{Caeruleuptychia} had a variation from
\texttt{r = 0.5234}, higher than the background net diversification
rate, to almost zero. This means that observed patterns and conclusions
can be completely contradictory depending on tree choice.

In this study, the effect of phylogenetic uncertainty on the inferred
diversification splits by MEDUSA is amplified because some Nymphalidae
taxa appear to be strongly affected by long-branch attraction artifacts
{[}27{]}. Thus, the Bayesian runs are expected to recover alternative
topologies on the posterior distribution of trees, resulting in low
support and posterior probability values for the nodes. For example,
posterior probability values for clades in Satyrini are very low
{[}10{]}. As a result, MEDUSA inferred a net diversification rate
increase for part of Satyrini in the maximum clade credibility tree, but
this was recovered only in 13\% of probabilities in the MultiMEDUSA
analysis on the random sample of trees.

If there is strong phylogenetic signal for increases or decreases in net
diversification rates for a node, it is expected that these splits would
be inferred by MEDUSA in most of the posterior distribution of trees.
However, weak phylogenetic signal for some nodes can cause some clades
to be absent in some trees and MEDUSA will be unable to estimate any
diversification shift (due to a non-existent node). This is the reason
why MEDUSA estimated net diversification rate splits with more than 90\%
of probabilities in the sample of trees for only four splits: the genus
\emph{Charaxes}, the genus \emph{Ypthima}, part of Ithomiini and the
clade \emph{Callicore} + \emph{Diaethria} (Fig. 3), while estimating
splits for other lineages with much lower probability.

The clade Ithomiini and the non-basal danaids are well supported by high
posterior probability values in Reference 10. Therefore our MEDUSA
analyses recovered an increase in net diversification rate with
probability higher than 90\% in the posterior distribution of trees
(Fig. 3).

\subsubsection{Hostplant use and diversification in
Nymphalidae}\label{hostplant-use-and-diversification-in-nymphalidae}

\paragraph{Ithomiini}\label{ithomiini}

Keith Brown suggested that feeding on Solanaceae was an important event
in the diversification of Ithomiini butterflies {[}42{]}. Ithomiini
butterflies are exclusively Neotropical and most species feed on
Solanaceae hostplants during larval stage {[}37{]}. Optimizations of the
evolution of hostplant use on phylogenies evidence a probable shift from
Apocynaceae to Solanaceae in the ancestor of the tribe {[}37,43{]}.
Reference 9 found that the Gamma statistics, a LTT plot of an Ithomiini
phylogeny and the fit of the density-dependent model of diversification
are consistent with a burst of diversification in Ithomiini following
the shift from Apocynaceae to Solanaceae.

We investigated whether the strong signal for an increase in net
diversification rate for Ithomiini (found by MEDUSA) can be explained
due to the use of Solanaceae plants as hosts during larval stage. For
this, we used a Bayesian approach {[}44{]} to test whether the trait
``feeding on Solanaceae'' had any effect on the diversification of the
group.

Our BiSSE analysis, extended to take into account missing taxa and
phylogenetic uncertainty, shows a significantly higher net
diversification rate for Ithomiini taxa, which can be attributed to the
trait ``feeding on Solanaceae hostplants'' (Fig. 4). This is in
agreement with the findings of Reference 9 using other statistical
methods. Due to the fact that Ithomiini are virtually the only
nymphalids using Solanaceae as hostplants, it is possible that the trait
responsible for a higher diversification of Ithomiini might not be the
hostplant character. As noted by Reference 36, the responsible trait
might be a codistributed character such as a trait related to the
ability to digest secondary metabolites.

Solanaceae plants contain chemical compounds and it has been suggested
that the high diversity of Ithomiini is consistent with the
``escape-and-radiate scenario'' due to a shift onto Solanaceae {[}9{]}
and radiation scenarios among chemically different lineages of
Solanaceae plants {[}37,42{]}. According to this theory, the shift from
Apocynaceae to Solanaceae allowed Ithomiini to invade newly available
resources due to a possible key innovation that allowed them cope with
secondary metabolites of the new hosts. Additional studies are needed to
identify the actual enzymes that Ithomiini species might be using for
detoxification of ingested food as they have been found in other
butterfly groups {[}45{]}.

The increase in diversification inferred by MEDUSA ocurred after the
probable shift from Apocynaceae to Solanaceae, as the Solanaceae feeders
in the subtribes Melinaeina and Mechanitina are not included in the
diversification shift (shift number 5 in Fig. 1). The apparent
conflicting results from MEDUSA and BiSSE can be explained by the low
species-richness of the subtribes Melinaeina and Mechanitina compared to
the other subtribes included in the shift (52 versus 272 species). It
can be that MEDUSA is more conservative than BiSSE and is not including
Melinaeina and Mechanitina in the shift due to low species numbers.

Although the Solanaceae genera used by the Ithomiini clades are well
known {[}37{]}, we do not have any understanding on the physiological
routes involved in the detoxification of Solanaceae compounds by the
several lineages of Ithomiini. We can speculate that older lineages
exploiting a novel toxic resource {[}10,37{]} might not be too efficient
in metabolizing plant toxins and that younger lineages are able to deal
with toxins more efficiently, so that host switching events within
Solanaceae are possible, which can lead to higher diversification.
Studies in \emph{Papilio} species have reported that detoxification
enzymes can become more efficient in metabolizing toxins than ancestral
configurations of the proteins, providing more opportunities for
hostplant switch {[}46{]}. This might be the reason why the basal
Ithomiini subtribes Melinaeina and Mechanitina are so species-poor and
restricted to few Solanaceae hosts {[}37{]}, while recent subtribes are
species-rich and have expanded their host range into several Solanaceae
lineages {[}37{]}. It might be that the switch to feeding in Solanaceae
was an important event in the evolutionary history of Ithomiini, but the
actual radiation occurred after critical physiological changes (a
probable key innovation) allowed efficient detoxification of Solanaceae
toxins.

The diffuse cospeciation hypothesis predicts almost identical ages of
insects and their hostplants, while the ``resource abundance-dependent
diversity'' and the ``escape-and-radiate'' hypotheses state that insects
diversify after their hostplants {[}2--4{]}. Reference 45 found strong
evidence for a model of speciation congruent with Ehrlich and Raven's
hypothesis in Pieridae butterflies due to, in addition to the
identification of a key innovation, a burst of diversification in
glucosinolate-feeding taxa shortly afterwards (with a lag of
\textasciitilde{}10 MY). According to a recent dated phylogeny of the
Angiosperms {[}47{]}, the family Solanaceae split from its sister group
about 59 (49-68) MYA and diversification started (crown group age)
around 37 (29-47) MYA. Reference 10 give the corresponding ages for
Ithomiini as 45 (39-53) and 37 (32-43) MYA, respectively. Thus, current
evidence shows that Solanaceae and Ithomiini might have diversified
around the same time, during the Late Eocene and Oligocene, and this
would be congruent with the diffuse cospeciation hypothesis.

The uplift of the Andes was a major tectonic event that underwent higher
activity during the Oligocene {[}48{]}. This caused climatic changes in
the region that affected the flora and fauna of the time, which
coincides with the diversification of modern montane plant and animal
taxa {[}49{]} including Ithomiini butterflies and Solanaceae hostplants.
Moreover, all Solanaceae clades currently present in New World
originated in South America {[}50{]} as well as Ithomiini butterflies
{[}10{]}. This suggests a process of ``diffuse cospeciation'' of
Ithomiini and hostplants.

\paragraph{Danaini}\label{danaini}

Our MultiMEDUSA approach gave a significant slowdown in net
diversification rate in the subtribe Danaina of the Danini. Both Danaina
and the sister clade Euploeina feed mainly on Apocynaceae and thus a
hostplant shift should not be responsible for the observed slowdown of
diversification in the Danaina. As expected, our BiSSE analysis of
Apocynaceae feeders shows that there is no effect of feeding on this
plant family on the net diversification rates of Nymphalidae lineages.
Many of the Danaina are large, strong fliers, highly migratory and
involved in mimicry rings. Among them is for example the monarch
(\emph{Danaus plexippus}), probably the most well known of all migratory
butterflies. The causes for a lower net diversification rate in the
Danaina remains to be investigated, but their great dispersal power
might be involved in preventing allopatric speciation. It has been found
in highly vagile species in the nymphalid genus \emph{Vanessa} that
dispersal has homogenized populations due to gene flow, as old and
vagile species seem to be genetically homogeneous while younger
widespread species show higher genetic differentiation in their
populations {[}51{]}.

\paragraph{\emph{Charaxes}}\label{charaxes}

The genus \emph{Charaxes} contain 193 species distributed in the Old
World with highest diversity in the Afrotropical region. These
butterflies are also very strong fliers, but contrary to Danaina, which
are specialized Apocynaceae feeders, \emph{Charaxes} are known to feed
on at least 28 plant families in 18 orders {[}29{]} and some species
appear to be polyphagous {[}52{]}. Reference 53 showed that most of the
diversification of the genus occurred from the late Oligocene to the
Miocene when there were drastic global climatic fluctuations, indicating
that the diversity mainly is driven by climate change. Reference 52
found that climatic changes during the Pliocene, Pleistocene as well as
dispersal and vicariance might have been responsible for the high
diversification of the genus.

\paragraph{Satyrini}\label{satyrini}

The diverse tribe Satyrini radiated simultaneously with the radiation of
their main hostplant, grasses, during the climatic cooling in the
Oligocene {[}11{]}. Thus, it is somewhat surprising that part of
Satyrini (the subtribes Euptychiina, Satyrina and Pronophilina) were
found to have accelerated diversification in only 13\% of probabilities
in the trees from the posterior distribution. Although this can be
attributed to low phylogenetic signal {[}10{]}, the clade Satyrini is
very robust {[}10{]} and MEDUSA failed to identify any significant
accelerated net diversification rate for Satyrini. It appears that the
radiation of Satyrini as a whole was not remarkably fast and therefore
not picked up by MEDUSA.

The origin of the tribe Satyrini is not completely clear (originated
either in the Neotropical or Eastern Palaearctic, Oriental and/or
Indo-Australian regions {[}27{]} and their radiation involved colonizing
almost all continents starting from their place of origin. For
butterflies, a prerequisite for colonizing new areas is that suitable
hostplants are already present. Therefore, it should be expected that
the diversification of Satyrinae occurred in a stepwise manner, with
pulses or bursts of diversification for certain lineages but unlikely
for the tribe Satyrini as a whole.

We found that even though MEDUSA estimated several diversification
shifts in the maximum clade credibility tree of Nymphalidae, only a few
of these splits were found in more than 90\% of the trees from the
posterior distribution. In the literature, it is common practice that
conclusions are based on the splits estimated on the maximum clade
credibility tree. However, by using a MultiMEDUSA approach, we found
that some of this splits might be greatly affected by phylogenetic
uncertainty. Moreover, some of these splits can be recovered either as
increases or decreases in net diversification rate depending on the tree
from the posterior distribution that was used for analysis. This means
that contradictory conclusions would be made if only the maximum clade
credibility tree was used for analysis. MEDUSA appears to be sensitive
to the number of nodes with high posterior probability and width of age
confidence intervals. For our data, it would be necessary to obtain a
posterior distribution of trees with no conflicting topology, and very
similar estimated ages for nodes in order to consistently recover most
of the diversification splits on the posterior distribution of trees
that were inferred by MEDUSA on the MCC tree.

Our MultiMEDUSA approach to perform analyses on the posterior
distribution of trees found strong support for an increase in net
diversification rate in the tribe Ithomiini and the genus
\emph{Charaxes}, and for a decrease in net diversification rate in the
subtribe Danaina. Due to phylogenetic uncertainty, we did not obtain
strong support for other diversification splits in Nymphalidae. Our
BiSSE analysis corroborated other studies in that the trait ``feeding on
Solanaceae'', or a codistributed character, was important in the
diversification of Ithomiini butterflies. However, by applying MEDUSA we
found that a critical character in the radiation of the group might have
appeared after the shift from Apocynaceae to Solanaceae. We also found
that the trait ``feeding on Apocynaceae'' is not responsible for the
slowdown of diversification in Danaina. Ithomiini and Solanaceae
diversified near simultaneously, which is in agreement with the diffuse
co-speciation hypothesis {[}2,4{]}.

\subsection{Acknowledgments}\label{acknowledgments}

We are thankful to Mark Cornwall for help with the script to extend
MEDUSA to include phylogenetic uncertainty, Niklas Wahlberg for
commenting on the manuscript and giving us the posterior distribution of
trees, Luke Harmon for commenting on the manuscript and anonymous
reviewers for their comments, which greatly improved the manuscript,
Jessica Slove Davidson and Niklas Janz for access to their hostplant
data. The study was supported by a Kone Foundation grant (awarded to
Niklas Wahlberg), Finland (C. Peña) and the Research Council of Norway
(grant no. 204308 to M. Espeland). We acknowledge CSC--IT Center for
Science Ltd. (Finland) for the allocation of computational resources.

\subsection{Figure legends}\label{figure-legends}

\texttt{Figure 1.} Results of the MEDUSA analysis run on the maximum
clade credibility tree from Reference 10. Rate shifts were estimated for
the following nodes (besides the background rate): 2) Limenitidinae +
Heliconiinae, 3) \emph{Ypthima}, 4) \emph{Charaxes}, 5) Ithomiini in
part, 6) Satyrini, 7) Coenonymphina 8) Phyciodina in part, 9) Danaini in
part, 10) \emph{Coenonympha}, 11) \emph{Caeruleuptychia} +
\emph{Magneuptychia}, 12) \emph{Callicore} + \emph{Diaethria}, 13)
Satyrina, 14) Mycalesinaa, 15) \emph{Pedaliodes}, 16) \emph{Dryas} +
\emph{Dryadula}, 17) \emph{Taenaris}, 18) Pseudergolinae.

\texttt{Figure 2.} Boxplot of the range of diversification values for
tips or clades estimated by MEDUSA on the 1000 random trees from the
posterior distribution of the Nymphalidae phylogeny. The tips or clades
shown are those present on the maximum clade credibility tree from
Reference 10.

\texttt{Figure 3.} Results of the MultiMEDUSA analysis on 1000 randomly
sampled trees from the posterior distribution of the Nymphalidae
phylogeny. Bars show the probability for nodes of being recovered as
significant increases or decreases in net diversification rates by
MultiMEDUSA.

\texttt{Figure 4.} BiSSE analysis of diversification of nymphalids due
to feeding on Solanaceae hostplants. Speciation and net diversification
rates are significantly higher in Solanaceae feeders (\$lambda\$1, r1).

\subsection{Tables}\label{tables}

\texttt{Table 1.} Significant net diversification rate shifts found in
the MEDUSA analysis of the Nymphalid maximum clade credibility tree.
Split.Node = node number, Model = preferred diversification model by
MEDUSA, r = net diversification rate, LnLik.part = log likelihood value.

\texttt{Table 2.} Differences in rates estimated by MEDUSA on the MCC
tree from Wahlberg and the MultiMEDUSA approach on 1000 random trees
from the posterior distribution.

\texttt{Table 3.} Likelihood ratio test between the model of increased
diversification of nymphalids feeding on Solanaceae against a model
forcing equal speciation rates (no effect on diversification).

\section{Supp. mat.}\label{supp.-mat.}

\textbf{\texttt{supp\_mat\_01\_genus.nex}}: MCC Nymphalidae tree from
Reference 10.

\textbf{\texttt{supp\_mat\_02\_run\_medusa\_on\_mct.R}}: Run MEDUSA on
MCC tree from Reference 10.

\textbf{\texttt{supp\_mat\_03\_richness.nex}}: Species richness for
lineages in Nymphalidae.

\textbf{\texttt{supp\_mat\_04\_MEDUSA\_results.txt}}: Results of running
MEDUSA on the MCC tree.

\textbf{\texttt{supp\_mat\_05\_1000\_random\_trees\_no\_outgroups.nex}}:
1000 random trees from Reference 10.

\textbf{\texttt{supp\_mat\_06\_run\_multimedusa.R}}: Run MultiMEDUSA on
1000 random trees from Reference 10.

\textbf{\texttt{supp\_mat\_07\_multimedusa\_on\_1000\_trees.txt}}: Raw
results from the MultiMEDUSA run on the random sample of trees from the
posterior distribution.

\textbf{\texttt{supp\_mat\_08\_multimedusa\_result01.pdf}}: Summary of
MulitMEDUSA analysis on the 1000 trees from the posterior distribution.

\textbf{\texttt{supp\_mat\_09\_mcc\_tree\_from\_random\_sample\_of\_1000.tree}}:
MCC tree from the 1000 random trees selected from the posterior
distribution.

\textbf{\texttt{supp\_mat\_10\_run\_medusa\_on\_mct\_from\_1000\_trees.R}}:
Run MEDUSA on MCC tree from the 1000 random trees selected from the
posterior distribution.

\textbf{\texttt{supp\_mat\_11\_results\_of\_MEDUSA\_run\_on\_mcc\_from\_1000\_trees.txt}}:
Summary results from a MEDUSA run on the MCC tree from the 1000 random
trees selected from the posterior distribution.

\textbf{\texttt{supp\_mat\_12\_MEDUSA\_mcc\_tree\_from\_random\_sample\_of\_1000\_trees.pdf}}:
Figure for MEDUSA run on MCC tree from random 1000 trees.

\textbf{\texttt{supp\_mat\_13\_results\_of\_Multi-MEDUSA\_run\_on\_1000\_trees.txt}}:
Summary results from a MultiMEDUSA run on the 1000 random trees selected
from the posterior distribution.

\textbf{\texttt{supp\_mat\_14\_multimedusa\_result\_prob\_nodes.pdf}}:
Probability of nodes with estimated rates from a MultiMEDUSA run on the
1000 random trees selected from the posterior distribution.

\textbf{\texttt{supp\_mat\_15\_host\_plant\_data.csv}}: Hostplants of
Nymphalidae butterflies recorded from the literature.

\textbf{\texttt{supp\_mat\_16\_host\_plant\_data\_references.csv}}:
References for hostplants data.

\textbf{\texttt{supp\_mat\_17\_hostplant\_states.csv}}: Data matrix with
character states for hosplant use.

\textbf{\texttt{supp\_mat\_18\_bisse\_source.R}}: R code for running the
BiSSE analysis.

\textbf{\texttt{supp\_mat\_19\_bisse\_mcmc\_run.csv}}: Raw results for
the BiSSE analysis.

\textbf{\texttt{supp\_mat\_20\_combined\_bisse.csv}}: Raw results for
the combined BiSSE analysis.

\texttt{Figure S1.} Boxplot of speciation (lambda) and extinction (mu)
values for the 95\% credibility intervals of values estimated by BiSSE
analysis of diversification due to feeding on Solanaceae plants.

\texttt{Figure S2.} BiSSE analysis of diversification of nymphalids due
to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
\emph{Hypanartia} as non-Solanaceae feeders. The same pattern is
recovered, speciation and net diversification rates are significantly
higher for Solanaceae feeders (\$lambda\$1, r1).

\texttt{Figure S3.} Net diversification rates of nymphalids feeding on
Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
the 1000 trees from the posterior distribution.

\texttt{Figure S4.} Boxplot of speciation and extinction values for the
95\% credibility intervals of values estimated by BiSSE analysis of
diversification due to feeding on Solanaceae plants on the combined
post-burnin runs on 1000 trees from the posterior distribution.

\texttt{Figure S5.} BiSSE analysis of diversification of nymphalids due
to feeding on Apocynaceae hostplants. Speciation and net diversification
rates are similar.

\subsection{References}\label{references}

1. Mitter C, Farrell B, Wiegmann B (1988) The phylogenetic study of
adaptive zones: has phytophagy promoted insect diversification? Am Nat
132: 107--128.
doi:\href{http://dx.doi.org/10.1086/284840}{10.1086/284840}.

2. Nyman T, Linder HP, Peña C, Malm T, Wahlberg N (2012) Climate-driven
diversity dynamics in plants and plant-feeding insects. Ecol Lett 15:
889--898. Available:
\url{http://doi.wiley.com/10.1111/j.1461-0248.2012.01782.x}. Accessed 29
March 2013.

3. Ehrlich PR, Raven PH (1964) Butterflies and plants: a study in
coevolution. Evolution 18: 586--608.

4. Janz N (2011) Ehrlich and raven revisited: mechanisms underlying
codiversification of plants and enemies. Annu Rev Ecol Evol S 42:
71--89. Available:
\url{http://www.annualreviews.org/doi/abs/10.1146/annurev-ecolsys-102710-145024}.

5. Nylin S, Slove J, Janz N (2013) Host plant utilization, host range
oscillations and diversification in nymphalid butterflies: a
phylogenetic investigation. Evolution: in press.

6. Nieukerken EJ van, Kaila L, Kitching IJ, Kristensen NP, Lees DC, et
al. (2011) Order lepidoptera. In:. Animal biodiversity: an outline of
higher-level classification and survey of taxonomic richness. Zootaxa.
pp. 212--221.

7. Heikkilä M, Kaila L, Mutanen M, Peña C, Wahlberg N (2012) Cretaceous
origin and repeated tertiary diversification of the redefined
butterflies. P Roy Soc B 279: 1093--1099. Available:
\href{http://rspb.royalsocietypublishing.org/cgi/doi/10.1098/rspb.2011.1430 http://www.ncbi.nlm.nih.gov/pubmed/21920981}{http://rspb.royalsocietypublishing.org/cgi/doi/10.1098/rspb.2011.1430
http://www.ncbi.nlm.nih.gov/pubmed/21920981}.

8. Elias M, Joron M, Willmott KR, Silva-Brandão KL, Kaiser V, et al.
(2009) Out of the andes: patterns of diversification in clearwing
butterflies. Mol Ecol 18: 1716--1729. Available:
\url{http://www.ncbi.nlm.nih.gov/pubmed/19386035}.

9. Fordyce JA (2010) Host shifts and evolutionary radiations of
butterflies. P Roy Soc B 277: 3735--3743. Available:
\url{http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=2992698\&tool=pmcentrez\&rendertype=abstract}.

10. Wahlberg N, Leneveu J, Kodandaramaiah U, Peña C, Nylin S, et al.
(2009) Nymphalid butterflies diversify following near demise at the
Cretaceous/Tertiary boundary. P Roy Soc B 276: 4295--4302. Available:
\url{http://rspb.royalsocietypublishing.org/cgi/doi/10.1098/rspb.2009.1303}.
Accessed 29 March 2013.

11. Peña C, Wahlberg N (2008) Prehistorical climate change increased
diversification of a group of butterflies. Biol Lett 4: 274--278.
Available:
\url{http://rsbl.royalsocietypublishing.org/cgi/doi/10.1098/rsbl.2008.0062}.
Accessed 29 March 2013.

12. Ferrer-Paris JR, Sánchez-Mercado A, Viloria ÁL, Donaldson J (2013)
Congruence and diversity of butterfly-host plant associations at higher
taxonomic levels. PloS one 8: e63570.

13. Alfaro ME, Santini F, Brock C, Alamillo H, Dornburg A, et al. (2009)
Nine exceptional radiations plus high turnover explain species diversity
in jawed vertebrates. P Natl Acad Sci USA 106: 13410--13414. Available:
\url{http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=2715324\$/backslash\$\&tool=pmcentrez\$/backslash\$\&rendertype=abstract}.

14. Harmon LJ, Rabosky DL, FitzJohn RG, Brown JW (2011) turboMEDUSA:
MEDUSA: modeling evolutionary diversification using stepwise AIC.

15. FitzJohn RG (2012) Diversitree: comparative phylogenetic analyses of
diversification in r. Method Ecol Evol 3: 1084--1092. Available:
\url{http://doi.wiley.com/10.1111/j.2041-210X.2012.00234.x}.

16. Lamas G (2004) Checklist: part 4A. hesperioidea-papilionoidea.
Gainesville.

17. Matos-Maraví PF, Peña C, Willmott KR, Freitas AV, Wahlberg N (2013)
Systematics and evolutionary history of butterflies in the ``Taygetis
clade'' (nymphalidae: satyrinae: euptychiina): towards a better
understanding of neotropical biogeography. Mol Phylogenet Evol 66:
54--68. Available:
\url{http://linkinghub.elsevier.com/retrieve/pii/S1055790312003624}.
Accessed 29 March 2013.

18. Brower AVZ, Wahlberg N, Ogawa JR, Boppré M, Vane-Wright RI (2010)
Phylogenetic relationships among genera of danaine butterflies
(lepidoptera: nymphalidae) as implied by morphology and DNA sequences.
Syst Biodivers 8: 75--89.
doi:\href{http://dx.doi.org/10.1080/14772001003626814}{10.1080/14772001003626814}.

19. Kodandaramaiah U, Lees DC, Müller CJ, Torres E, Karanth KP, et al.
(2010) Phylogenetics and biogeography of a spectacular old world
radiation of butterflies: the subtribe mycalesina (lepidoptera:
nymphalidae: satyrini). BMC Evol Biol 10: 172. Available:
\url{http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=2898688\&tool=pmcentrez\&rendertype=abstract}.

20. Kodandaramaiah U, Peña C, Braby MF, Grund R, Müller CJ, et al.
(2010) Phylogenetics of coenonymphina (nymphalidae: satyrinae) and the
problem of rooting rapid radiations. Mol Phylogenet Evol 54: 386--394.
Available:
\url{http://linkinghub.elsevier.com/retrieve/pii/S1055790309003273}.
Accessed 29 March 2013.

21. Ortiz-Acevedo E, Willmott KR (2013) Molecular systematics of the
butterfly tribe preponini (nymphalidae: charaxinae). Syst Entomol:
no--o. Available: \url{http://doi.wiley.com/10.1111/syen.12008}.

22. De-Silva DL, Day JJ, Elias M, Willmott KR, Whinnett A, et al. (2010)
Molecular phylogenetics of the neotropical butterfly subtribe oleriina
(nymphalidae: danainae: ithomiini). Mol Phylogenet Evol 55: 1032--1041.
Available: \url{http://www.ncbi.nlm.nih.gov/pubmed/20079859}.

23. Freitas AVL, Brown Jr. KS (2004) Phylogeny of the nymphalidae
(lepidoptera). Syst Biol 53: 363--383. Available:
\url{http://sysbio.oxfordjournals.org/cgi/doi/10.1080/10635150490445670}.

24. Peña C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
(2006) Higher level phylogeny of satyrinae butterflies (lepidoptera:
nymphalidae) based on DNA sequence data. Mol Phylogenet Evol 40: 29--49.
Available: \url{http://www.ncbi.nlm.nih.gov/pubmed/16563805}.

25. Penz CM (1999) Higher level phylogeny for the passion-vine
butterflies (nymphalidae, heliconiinae) based on early stage and adult
morphology. Zool J Linn Soc 127: 277--344. Available:
\url{http://doi.wiley.com/10.1111/j.1096-3642.1999.tb00680.x}. Accessed
03 September 2013.

26. Silva-Brandão KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
tribe acraeini (lepidoptera, nymphalidae, heliconiinae) and the
evolution of host plant use. Mol Phylogenet Evol 46: 515--531.
Available: \url{http://www.ncbi.nlm.nih.gov/pubmed/18191590}.

27. Peña C, Nylin S, Wahlberg N (2011) The radiation of satyrini
butterflies (nymphalidae: satyrinae): a challenge for phylogenetic
methods. Zool J Linn Soc 161: 64--87. Available:
\url{http://doi.wiley.com/10.1111/j.1096-3642.2009.00627.x}. Accessed 29
March 2013.

28. Peña C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
history of the butterfly subtribe euptychiina (lepidoptera, nymphalidae,
satyrinae). Zool Scr 39: 243--258. Available:
\url{http://blackwell-synergy.com/doi/abs/10.1111/j.1463-6409.2010.00421.x}.

29. Ackery PR (1988) Hostplants and classification: a review of
nymphalid butterflies. Biol J Linn Soc 33: 95--203.

30. Dyer LA, Gentry GL (2002) Caterpillars and parasitoids of a tropical
lowland wet forest. Available: \url{http://www.caterpillars.org}.
Accessed 17 April 2013.

31. Beccaloni G, Viloria Á, Hall S, Robinson G (2008) Catalogue of the
hostplants of the neotropical butterflies. London: The Natural History
Museum.

32. Janzen DH, Hallwachs W (2009) Dynamic database for an inventory of
the macrocaterpillar fauna, and its food plants and parasitoids, of area
de conservacion guanacaste (aCG), northwestern costa rica. Available:
\url{http://janzen.sas.upenn.edu}. Accessed 17 April 2013.

33. Core Team R (2013) R: a language and environment for statistical
computing. Vienna, Austria: R Foundation for Statistical Computing.
Available: \url{http://www.r-project.org}.

34. Popescu A-A, Huber KT, Paradis E (2012) ape 3.0: new tools for
distance-based phylogenetics and evolutionary analysis in r.
Bioinformatics 28: 1536--1537.

35. Harmon LJ, Weir JT, Brock CD, Glor RE, Challenger W (2008) GEIGER:
investigating evolutionary radiations. Bioinformatics 24: 129--131.
Available: \url{http://www.ncbi.nlm.nih.gov/pubmed/18006550}.

36. Maddison WP, Midford PE, Otto SP (2007) Estimating a binary
character's effect on speciation and extinction. Syst Biol 56: 701--710.
Available: \url{http://www.ncbi.nlm.nih.gov/pubmed/17849325}.

37. Willmott KR, Freitas AVL (2006) Higher-level phylogeny of the
ithomiinae (lepidoptera: nymphalidae): classification, patterns of
larval hostplant colonization and diversification. Cladistics 22:
297--368. Available:
\url{http://doi.wiley.com/10.1111/j.1096-0031.2006.00108.x}.

38. Scott JA (1986) The butterflies of north america. Stanford
University Press.

39. Davis MP, Midford PE, Maddison W (2013) Exploring power and
parameter estimation of the BiSSE method for analyzing species
diversification. BMC Evol Biol 13: 38. Available:
\url{http://www.biomedcentral.com/1471-2148/13/38}.

40. Litman JR, Danforth BN, Eardley CD, Praz CJ (2011) Why do leafcutter
bees cut leaves? new insights into the early evolution of bees. P Roy
Soc B 278: 3593--3600. Available:
\url{http://www.pubmedcentral.nih.gov/articlerender.fcgi?artid=3189370\$/backslash\$\&tool=pmcentrez\$/backslash\$\&rendertype=abstract}.

41. Ryberg M, Matheny PB (2012) Asynchronous origins of ectomycorrhizal
clades of agaricales. P Roy Soc B 279: 2003--2011. Available:
\url{http://www.ncbi.nlm.nih.gov/pubmed/22171078}.

42. Brown Jr. KS (1987) Chemistry at the Solanaceae/Ithomiinae
interface. Ann Mo Bot Gard 74: 359--397.

43. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brandão KL, Whinnett A, et
al. (2006) Phylogenetic relationships among the ithomiini (lepidoptera:
nymphalidae) inferred from one mitochondrial and two nuclear gene
regions. Syst Entomol 31: 288--301. Available:
\url{http://doi.wiley.com/10.1111/j.1365-3113.2006.00321.x}.

44. FitzJohn RG, Maddison WP, Otto SP (2009) Estimating trait-dependent
speciation and extinction rates from incompletely resolved phylogenies.
Syst Biol 58: 595--611. Available:
\url{http://www.ncbi.nlm.nih.gov/pubmed/20525612}.

45. Wheat CW, Vogel H, Wittstock U, Braby MB, Underwood D, et al. (2007)
The genetic basis of a plant-insect coevolutionary key innovation. P
Natl Acad Sci USA 104: 20427--20431.
doi:\href{http://dx.doi.org/10.1073/pnas.0706229104}{10.1073/pnas.0706229104}.

46. Li W, Schuler MA, Berenbaum MR (2003) Diversification of
furanocoumarin-metabolizing cytochrome p450 monooxygenases in two
papilionids: specificity and substrate encounter rate. P Natl Acad Sci
USA 100: 14593--14598. Available:
\url{http://www.pnas.org/content/100/suppl.2/14593.abstract}.

47. Bell CD, Soltis DE, Soltis PS (2010) The age and diversification of
the angiosperms re-revisited. Am J Bot 97: 1296--1303. Available:
\url{http://www.ncbi.nlm.nih.gov/pubmed/21616882}.

48. Somoza R (1998) Updated nazca (farallon) - south america relative
motions during the last 40 my: implications for mountain building in the
central andean region. J S Am Earth Sci 11: 211--215.

49. Hoorn C, Wesselingh FP, Steege H ter, Bermudez MA, Mora A, et al.
(2010) Amazonia through time: andean uplift, climate change, landscape
evolution, and biodiversity. Science 330: 927--931. Available:
\url{http://www.ncbi.nlm.nih.gov/pubmed/21071659}.

50. Olmstead RG (2013) Phylogeny and biogeography in solanaceae,
verbenaceae and bignoniaceae: a comparison of continental and
intercontinental diversification patterns. Bot J Linn Soc 171: 80--102.
Available: \url{http://doi.wiley.com/10.1111/j.1095-8339.2012.01306.x}.

51. Wahlberg N, Rubinoff D (2011) Vagility across vanessa (lepidoptera:
nymphalidae): mobility in butterfly species does not inhibit the
formation and persistence of isolated sister taxa. Syst Entomol 36:
362--370. Available:
\url{http://doi.wiley.com/10.1111/j.1365-3113.2010.00566.x}. Accessed 16
April 2013.

52. Müller CJ, Wahlberg N, Beheregaray LB (2010) `After africa': the
evolutionary history and systematics of the genus charaxes ochsenheimer
(lepidoptera: nymphalidae) in the indo-pacific region. Biol J Linn Soc
100: 457--481. Available:
\url{http://doi.wiley.com/10.1111/j.1095-8312.2010.01426.x}.

53. Aduse-Poku K, Vingerhoedt E, Wahlberg N (2009) Out-of-africa again:
a phylogenetic hypothesis of the genus charaxes (lepidoptera:
nymphalidae) based on 5 gene regions. Mol Phylogenet Evol 53: 463--478.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2009.06.021}{10.1016/j.ympev.2009.06.021}.

\end{document}