### 2014-10-08 response_to_reviewers.tex
punto 2

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in,a4paper]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

\usepackage{color}
\date{}

\begin{document}

\section{Response to reviewers}\label{response-to-reviewers}

\subsection{Reviewer 1}\label{reviewer-1}

\textbf{1. Reviewer's Comment:}

Introduction 1. Second paragraph, second sentence: The fact that the
Nymphalids are the ``largest family'' of butterflies conveys an
exaggerated importance of their diversity based on idiosyncrasy of
Linnean taxonomy (the sister group to Nymphalidae, the clade that
includes riodinids and lycaenids, has over 6500 species). Consider
revising the sentence, omitting superlatives (``biggest'', ``largest'',
``most'', etc.).

\textbf{1. Response:}

\begin{quote}
\color{blue}
We have rewritten the sentence to: ``Nymphalidae contains around 6000
species {[}6{]}, and several members are considered model organisms in
evolutionary biology {[}7--9{]}.''
\end{quote}

\textbf{2. Reviewer's Comment:}

The authors should mention that the ideal approach would accommodate
uncertainty by sampling from a distribution of trees which reflected the
variation in confidence in various parts of the tree (such as a Bayesian
posterior distribution or a bootstrap sample). See Holder,M. and
Lewis,P.O. (2003) Phylogeny estimation: traditional and Bayesian
approaches. Nat. Rev.~Genet., 4, 275--284

\end{document}