### 2014-10-09 response_to_reviewers.md
punto 6

# Response to reviewers

## Reviewer 1

**1. Reviewer's Comment:**

Introduction
1. Second paragraph, second sentence: The fact that the Nymphalids are the “largest family” of butterflies conveys an exaggerated importance of their diversity based on idiosyncrasy of Linnean taxonomy (the sister group to Nymphalidae, the clade that includes riodinids and lycaenids, has over 6500 species). Consider revising the sentence, omitting superlatives (“biggest”, “largest”, “most”, etc.).

**1. Response:**

> We have rewritten the sentence to:
> "Nymphalidae contains around 6000 species [6], and several members are
considered model organisms in evolutionary biology [7–9]."

**2. Reviewer's Comment:**

The authors should mention that the ideal approach would accommodate uncertainty by sampling from a distribution of trees which reflected the variation in confidence in various parts of the tree (such as a Bayesian posterior distribution or a bootstrap sample). See Holder,M. and Lewis,P.O. (2003) Phylogeny estimation: traditional and Bayesian approaches. Nat. Rev. Genet., 4, 275–284

**2. Response:**

> We have specified that the sample of trees was taken from the Bayesian run in
Wahlberg et al., 2009.

**3. Reviewer's Comment:**

Middle of paragraph (and throughout the manuscript): revise the tree sampling to “...therefore randomly sampled 1000 trees from the posterior distribution”; in the remainder of the manuscript, refer to these as the “1000 sampled trees”, not “1000 random trees”, which can take on a very different meaning.

**3. Response:**

> We have corrected some sentences in the manuscript to specify that we are
referring to the same set of 1000 trees.

**4. Reviewer's Comment:**

Define what 'MuSSE' refers to (Multiple State Speciation and Extinction?).

**4. Response:**

> We have included the definition of MuSSE in the manuscript.

**5. Reviewer's Comment:**

A reference for the statement “most of Nymphalidae butterflies are restricted to use one plant family as hostplant” is warranted.

**5. Response:**

> We have included the numbers of single-family feeders based on calculations
of our own data on hostplant use across Nymphalidae genera. The text has been
modified accordingly:

> "Because most of Nymphalidae butterflies are restricted to use one plant
family as hostplant (232 genera use only one plant family versus 176 that use
        more than one; based on our data in supp. mat. 15), the ..."

**6. Reviewer's Comment:**

At the end of this paragraph, the authors indicated they tested the effect of character coding on their analyses. A more explicit description of the test would be useful, along with predictions of how results would be effected (i.e. how would a large effect vs. no effect of coding Hypanartia and Vanessa as Solanaceae feeders manifest in the results?).

**6. Response:**

> We included the following sentence to clarify the issue: **"We expected that
coding these two genera as either absence or presence would not have a
significant effect on our results."**