### 2014-10-13 response_to_reviewers.md
punto 16

# Response to reviewers

## Reviewer 1

**1. Reviewer's Comment:**

Introduction
1. Second paragraph, second sentence: The fact that the Nymphalids are the “largest family” of butterflies conveys an exaggerated importance of their diversity based on idiosyncrasy of Linnean taxonomy (the sister group to Nymphalidae, the clade that includes riodinids and lycaenids, has over 6500 species). Consider revising the sentence, omitting superlatives (“biggest”, “largest”, “most”, etc.).

**1. Response:**

> We have rewritten the sentence to:
> "Nymphalidae contains around 6000 species [6], and several members are
considered model organisms in evolutionary biology [7–9]."

**2. Reviewer's Comment:**

The authors should mention that the ideal approach would accommodate uncertainty by sampling from a distribution of trees which reflected the variation in confidence in various parts of the tree (such as a Bayesian posterior distribution or a bootstrap sample). See Holder,M. and Lewis,P.O. (2003) Phylogeny estimation: traditional and Bayesian approaches. Nat. Rev. Genet., 4, 275–284

**2. Response:**

> We have specified that the sample of trees was taken from the Bayesian run in
Wahlberg et al., 2009.

**3. Reviewer's Comment:**

Middle of paragraph (and throughout the manuscript): revise the tree sampling to “...therefore randomly sampled 1000 trees from the posterior distribution”; in the remainder of the manuscript, refer to these as the “1000 sampled trees”, not “1000 random trees”, which can take on a very different meaning.

**3. Response:**

> We have corrected some sentences in the manuscript to specify that we are
referring to the same set of 1000 trees.

**4. Reviewer's Comment:**

Define what 'MuSSE' refers to (Multiple State Speciation and Extinction?).

**4. Response:**

> We have included the definition of MuSSE in the manuscript.

**5. Reviewer's Comment:**

A reference for the statement “most of Nymphalidae butterflies are restricted to use one plant family as hostplant” is warranted.

**5. Response:**

> We have included the numbers of single-family feeders based on calculations
of our own data on hostplant use across Nymphalidae genera. The text has been
modified accordingly:

> "Because most of Nymphalidae butterflies are restricted to use one plant
family as hostplant (232 genera use only one plant family versus 176 that use
        more than one; based on our data in supp. mat. 15), the ..."

**6. Reviewer's Comment:**

At the end of this paragraph, the authors indicated they tested the effect of character coding on their analyses. A more explicit description of the test would be useful, along with predictions of how results would be effected (i.e. how would a large effect vs. no effect of coding Hypanartia and Vanessa as Solanaceae feeders manifest in the results?).

**6. Response:**

> We included the following sentence to clarify the issue: **"We expected that
coding these two genera as either absence or presence would not have a
significant effect on our results."**

**7, 8, 9, 10. Reviewer's Comment:**

7. First paragraph, second sentence: change the tense of “run” to “ran”.
8. Second paragraph, third sentence: replace “Thus” with “Furthermore”.
9. There appears to be an orphaned right-parenthesis at the end of the second paragraph.
10. Third paragraph, last sentence: The node number is not really relevant, so beginning of the sentence could be revised to “For example, the Charaxes + Polyura clade was found in only 256 trees...”

**7, 8, 9, 10. Response:**

> We made the suggested corrections.

**11. Reviewer's Comment:**

Fourth paragraph: The point of the paragraph is not clear. What rates are being
compared? MEDUSA vs. MultiMEDUSA? Consider revising to make the point clearer
or omit the sentence.

**11. Response:**

> We have rewritten the sentence and now is more clear: **"For the diversification shifts found in both the MCC tree and most of the
samples of 1000 trees (frequency more than 90%; Table 2), the MultiMEDUSA
approach recovered different rates of diversification than
those found using when the MCC tree alone."**.

**12. Reviewer's Comment:**

Fifth paragraph, first sentence (and potentially elsewhere): Consider using a consistent means of presenting probability values. In this sentence, the authors refer to “90% of probability in the trees”, but in the preceding paragraph, the authors present probabilities as a decimal (“probability of being recovered higher than 0.90”). Granted, these two probabilities refer to different events, but perhaps it would be best to refer to them in a consistent manner, such as “0.90 posterior probability”.

**12. Response:**

> We have corrected this and normalized the probability values to decimals in the manuscript.

**13. Reviewer's Comment:**
The third sentence (length of analysis and burnin details) should be moved to the methods section.

**13. Response:**

> We made the suggested corrections.

**14. Reviewer's Comment:**

Mid-way through paragraph, discussion of likelihood ratio test: The source of this P-value needs to be described (chi-squared approximation with 1 degree of freedom?), here or in the methods section (see p. 617 in Pagel 1999 Syst. Biol. 48:612-622).

**14. Response:**

> We have include the statistics suggested by the reviewer: 
"**$\chi^2 = 12.3; 1 df; p < 0.001$**".

**15. Reviewer's Comment:**

The closing three sentences (about BiSSE performance) should be moved to the methods section.

**15. Response:**

> We made the suggested corrections.

**16. Reviewer's Comment:**

First paragraph: The authors' assessment of analyzing diversification shifts using a single tree, without accommodating our level of certainty in relationships, is justified; however, MEDUSA itself, that is, the approach using stepwise AIC to identify shifts in diversification rate, should not be the target of criticism. After all, MultiMEDUSA is doing the same thing, but on a sample of trees. Revise this paragraph with a more appropriate target for the criticism (i.e. analyses performed on single trees).

**16. Response:**

> We have reworded the paragraph to avoid blaming the method:

> "The MEDUSA method has been used to infer changes in net diversification rates in a phylogenetic
tree. Since its publication [16] the results of using MEDUSA on a single tree, the maximum clade
credibility tree, have been used for generation of hypotheses and discussion [10,42,43]. However, different
diversification shifts and different rates of diversification are found for certain lineages when phylogenetic
uncertainty was taken into account by using MEDUSA on a random sample of trees from the posterior
distribution of a Bayesian run. We found that some diversification splits, estimated on the Nymphalidae
maximum clade credibility tree, were found with very low probability in the random sample of 1000 trees
from the posterior distribution (Fig. 3, Table 2). We also found that, even though the analyses estimated
the same diversification splits on two or more trees, the estimated net diversification rates could vary
widely (Fig. 2)."